---
title: "Simpler covariance exploration"
output:
  html_document: default
  html_notebook: default
---

```{r}
knitr::opts_chunk$set(message = FALSE)
knitr::opts_chunk$set(warning = FALSE)
```

```{r}
library(tidyverse)
library(cowplot)
theme_set(theme_cowplot())
library(roll)
```

#### Invent some TPCs (purple is f(T), orange is g(T))
```{r}
f_curve <- function(x) 0.02*x + 1 
g_curve <- function(x) 0.07*x 
p <- ggplot(data = data.frame(x = 0), mapping = aes(x = x)) 

tpcs <- p + 
	stat_function(fun = f_curve, color = "purple", size = 2) +
	stat_function(fun = g_curve, color = "orange", size = 2) +
	xlim(0, 35) + ylim(-0.1, 2) + 
	xlab("Temperature (°C)") + ylab("Trait") + geom_hline(yintercept = 0)

tpcs
```



#### make up temperature time series
```{r}
set.seed(101)
time <- seq_along(1:100)
temps <- data.frame(temperature = runif(100, min = 5, max = 33), time = time)

```



#### plot temps over time

```{r}
 temps %>% 
	ggplot(aes(x = time, y = temperature)) + geom_line() +
	ylab("Temperature (°C)") + xlab("Time (days)")
```


#### get trait values

```{r}
f_time_series <- data.frame(time = time, ftrait = sapply(X = temps$temperature, FUN = f_curve), temperature = temps$temperature)
g_time_series <- data.frame(time = time, gtrait = sapply(X = temps$temperature, FUN = g_curve), temperature = temps$temperature)
```

#### plot traits over time
```{r}
both <- left_join(f_time_series, g_time_series) %>% 
	mutate(R_0 = ftrait*gtrait)


both %>% 
	gather(key = trait_type, value = trait_value, ftrait, gtrait) %>% 
	mutate(trait_type = ifelse(trait_type == "ftrait", "f(T)", "g(T)")) %>% 
	ggplot(aes(x = time, y = trait_value, color = trait_type)) + geom_line() +
	scale_color_manual(values = c("purple", "orange")) +ylab("Trait value") + xlab("Time (days)") +
	theme(legend.position = "top")

```




Find the averages at different time slices: 2, 4, 5, 10, 20, 25, 50, 100. 
Here each panel corresponds to a given time slice length
```{r}
x <- 20

time_chunk <- function(x){

group <-  sort(rep(seq(from = 1, to = x), times = 100/x))

all <- both %>% 
	mutate(group = group) %>% 
	mutate(slice_length = 100/x) %>% 
	mutate(R_0 = ftrait*gtrait) %>% 
	group_by(group) %>% 
	mutate(`<f(T)g(T)>` = mean(R_0)) %>% 
	mutate(`<f(T)>`= mean(ftrait)) %>% 
	mutate(`<g(T)>` = mean(gtrait)) %>% 
	mutate(`Cov[f(T), g(T)]` = cov(ftrait, gtrait)) %>% 
	mutate(`<f(T)><g(T)> + Cov[f(T), g(T)]` = (`<f(T)>` * `<g(T)>`) + `Cov[f(T), g(T)]`) %>% View
	gather(key = estimate, value = value, 8:12)
}


x <-c(1, 2, 4, 5, 10, 20, 25, 50)

all_slices <- x %>% 
	map_df(.f = time_chunk)

all_slices %>% 
	ggplot(aes(x = time, y = value, color = estimate)) + geom_line() +
	facet_wrap( ~ slice_length)
```

Now plot the differences between `<f(T)><g(T)> + Cov[f(T), g(T)]` and `<f(T)g(T)>` (should be 0)
```{r}
all_slices %>% 
	spread(key = estimate, value = value) %>% 
	mutate(difference = `<f(T)><g(T)> + Cov[f(T), g(T)]` - `<f(T)g(T)>`) %>% 
	ggplot(aes(x = time, y = difference, color = factor(slice_length), group = slice_length)) + geom_line(size = 1) +
	scale_color_viridis_d(name = "Time slice length") + ylab("Difference \n <f(T)><g(T)> + Cov[f(T), g(T)] - <f(T)g(T)>") + xlab("Time")
```

Now get the average differences between `<f(T)><g(T)> + Cov[f(T), g(T)]` and `<f(T)g(T)>`
```{r}
all_slices %>% 
	spread(key = estimate, value = value) %>%  
	mutate(difference = `<f(T)><g(T)> + Cov[f(T), g(T)]` - `<f(T)g(T)>`) %>% 
	group_by(slice_length) %>% 
	summarise(average_difference = mean(difference)) %>% 
	ggplot(aes(x  = slice_length, y = average_difference)) + geom_point() +
	ylab("Average difference between \n `<f(T)><g(T)> + Cov[f(T), g(T)]` and `<f(T)g(T)>`") +
	xlab("Time slice length")
```

Now get  difference between `<f(T)><g(T)> + Cov[f(T), g(T)]` and `<f(T)g(T)>` using the whole time series
```{r}
difference <- both %>% 
	mutate(R_0 = ftrait*gtrait) %>% 
	summarise(`<f(T)g(T)>` = mean(R_0),
			  `<f(T)>`= mean(ftrait),
			  `<g(T)>` = mean(gtrait),
			  `Cov[f(T), g(T)]` = cov(ftrait, gtrait)) %>% 
	mutate(`<f(T)><g(T)> + Cov[f(T), g(T)]` = (`<f(T)>` * `<g(T)>`) + `Cov[f(T), g(T)]`) %>%
	mutate(difference = `<f(T)><g(T)> + Cov[f(T), g(T)]` - `<f(T)g(T)>`) 

difference
```
