---
title: "Fluc_estimation"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
NOTE: FOr reasons I don't understand lines below need to be run mulitple times before they make the changes in the dataframes...
109: component.estimation[[6]][component.estimation[[6]]$hr == 7, "X16.dtr9"] <- 0
110: component.estimation[[6]][component.estimation[[6]]$hr == 21, "X16.dtr12"] <- 0
111: component.estimation[[6]][component.estimation[[6]]$hr == 22, "X16.dtr9"] <- 0



#Rate Summation

##Major question;
Does rate summation breakdown? 

###Analysis pipeline
1. Import constant temperature thermal performance curves (component & R0)
      i. component: bite rate, EFD , lifetime egg production, lifespan, Mu; estimation via lifetime averages | 1st GC
      ii. R0 of standard vs. reformulated
2. Estimate R0 under fluctuating conditions using R0 from constant temperature (i. standard, Mord method, Shap base & ii. reformulated, lifetime)
3. Estimate individual trait performance under fluctuating conditions using constant temperature TPC curves from either i. First GC (or Mu.exp method) or ii. lifetime temp values.
3b. Use these component traits to parameritize R0.
4. Compare R0 generated from method A to method B.
5. Plot observed performance under fluctuation
6. Estimate R0 (standard) using measured values under fluctuation (i. 1st GC (Mu as exp), or ii. lifetime temp values).
7. Find a way to statistically compare differences between constant, expected fluc, and observed fluc.

#Temperature Profile under fluctuation
```{r}
library(reshape2)
library(tidyverse)

#create a vector of all daily temperature values
temp.programs <- read.csv("Data/temp.programs.csv")
#order dataframe by hr
temp.programs <- temp.programs[order(temp.programs$hr),]
#remove the 0.5 hr temps, we only need hourly intervals
   #create sequence of row to remove (everyother)
remove.r <- seq(2,length(temp.programs$hr-1), 2)
temp.programs<- temp.programs[-remove.r,]

#In case I need the hr later
hr <- temp.programs$hr

#Place into a list for apply statements later
temp.programs.list <-  list( temp_program = temp.programs[2:length(temp.programs)])  #removes hr, creates a 1 unit list of the dataframe containing all of the hourly temperatures for both deters.
```


#Estimating component trait values at fluctuation using TPCs from constant temperature (lifetime)
```{r}
library(ggplot2)
library(easynls)

#Import functions defined in TPC overlay, generated in AIC curve fits
#NOTE:These were added 1/12/18, check later for final version
bite.life_fun <- function(Temp){ 0.0001425*Temp*(Temp-(-2.5761438))*(40.1676495-Temp)^(0.5) }
#Note this is lifetime (#bites on dayx divided by #females alive each day averaged across each day females are alive)
EFD.life_fun <- function(Temp){ 0.008925*Temp*(Temp-(10.494189))*(36.546948 -Temp)^(0.5)}   
#Define function for P
Miaz_P_Quad_TPC.exp <- function(Temp){.5989 + 0.0333*Temp + (-0.0008)*Temp^2}
#Convert to Mu
Mu.exp_fun <- function(Temp){-log(Miaz_P_Quad_TPC.exp(Temp))}

life.expect_fun <-function(Temp){-.1183*Temp^2 + 4.4620*Temp -1.6611}
life.eggs_fun <-function(Temp){(-1233.9070)+125.4588*Temp + (-2.4786)*Temp^2}


#define other functions
Mord_P<-function(x){-0.000828*x^2 + 0.0367*x + 0.522}
Mord_Mortality <- function(x) {-log(Mord_P(x=x))}
Mord_EFD<-function(x){-0.153*x^2 + 8.61*x -97.7}
Mord_Bx <- function(x){Mord_EFD(x=x) / Mord_Mortality(x=x)}
#Mord_VectorComp <- function(x){-0.54*x^2 + 25.2*x -206}
#Mord_PDR <- function(x){0.000111* x *(x - 14.7)*(34.4 -x)^(0.5)}
Mord_MDR <- function(x){0.000111 *x *(x - 14.7)*(34 -x)^(0.5)}
Mord_PEA <- function(x){-0.00924*x^2 + 0.453*x -4.77}

Shapiro_PDR <- function(x){ .1574* log(x) - 0.4153 }
Shapiro_VectorComp <- function(x){-0.0032*x^2 + .1604 *x -1.105}

r0_constant <- function (x){(bite.life_fun(x)^2 * Shapiro_VectorComp(x)* exp(-1/(Shapiro_PDR(x)*life.expect_fun(x))) * life.eggs_fun(x)* Mord_PEA(x) * Mord_MDR(x) * life.expect_fun(x)^2) }
curve(r0_constant, from = 15, to = 40)


#Create a dataframe of y values for Miaz_a & Miaz_a_Mord_1stGCdur using the dataframe in list[1]; assign the output to list[2]
component.estimation <- list(bite.lifetime = as.data.frame(apply(temp.programs.list[[1]], 2, bite.life_fun)), 
                             EFD.lifetime = as.data.frame(apply(temp.programs.list[[1]], 2, EFD.life_fun)),
                             Mu.exp = as.data.frame(apply(temp.programs.list[[1]],2, Mu.exp_fun)),
                             life.expect = as.data.frame(apply(temp.programs.list[[1]], 2, life.expect_fun)),
                             life.eggs = as.data.frame(apply(temp.programs.list[[1]], 2, life.eggs_fun)),
                             r0 = as.data.frame(apply(temp.programs.list[[1]],2, r0_constant))) #store this result into a new list which will correspond with the time 

#Replace all Na values with 0 withing the outputs; and any negative values with 0
component.estimation[[2]][is.na(component.estimation[[2]])] <- 0  #same as component.estimation[["EFD.lifetime"]][is.na(component.estimation[[EFD.lifetime]])] <- 0 
component.estimation[[3]][component.estimation[[3]] < 0 ] <- 0
component.estimation[[4]][component.estimation[[4]] < 0 ] <- 0
component.estimation[[5]][component.estimation[[5]] < 0 ] <- 0

component.estimation[[6]][is.na(component.estimation[[6]])] <- 0 
component.estimation[[6]][component.estimation[[6]] < 0 ] <- 0

#NOTE; THERE IS A VERY WIERD VALUE IN THE DATA FRAME i'M REMOVING IT AND REPLACING IT WITH 0c
#NEED TO RUN THIS PART TWICE?
component.estimation[[6]][component.estimation[[6]]$hr == 7, "X16.dtr9"] <- 0
component.estimation[[6]][component.estimation[[6]]$hr == 21, "X16.dtr12"] <- 0
component.estimation[[6]][component.estimation[[6]]$hr == 22, "X16.dtr9"] <- 0

#Calculates column means in the dataframe
trait.estimate = lapply(component.estimation, function(x) {
  means = colMeans(x)
  means = as.data.frame(means)
  means$treatment = row.names(means)
  return(means)
})

names <- names(component.estimation)
names(trait.estimate) <- names

#Organizes mean trait.estimate values for easy plotting
trait.estimate = lapply(trait.estimate, function(x) {  #Applies this function to each element of x (id each dataframe)
  x$temp =  as.numeric( substr(x$treatment, 2, 3) ) #new columns based on sub strings of the previous name
  x$dtr = as.factor(substr(x$treatment,8,9))
    #remove row titled 'hr'
  row.names(x) <- seq(1:nrow(x))
  x = x[!x$treatment == "hr",]
    return(x)
})


#Then plot all of these again from above with the function of the that plot included

#Make a plot of how they change with time; a seperate plot for each dataframe; 1. add hr column to each df, 2. nested inner loop to create a seperate plot for each trait, with a series corresponding for each TEMP*DTR combo

plots_hourly <- list()
for(i in 1:length(component.estimation)) {  #outter loop start
  component.estimation[[i]]$hr <- hr #add hr vector to end of dataframe
    
  temporary.df <- melt(component.estimation[[i]] ,  id.vars = 'hr', variable.name = 'treatment')
  
  plots_hourly[[i]] <- ggplot(temporary.df, aes(hr , value), group = treatment, colour = treatment, fill = treatment) +
      geom_point(aes(colour = treatment, fill = treatment))+
      geom_line(aes(colour= treatment, linetype = treatment))+
      ggtitle( names(component.estimation)[i])+  #title to correspond to trait
      theme_classic()
  
  } #outter loop close

names(plots_hourly) <- names

#Look at hourly estimate plots
plots_hourly[[1]]
plots_hourly[[2]]
plots_hourly[[3]]
plots_hourly[[4]]
plots_hourly[[5]]
plots_hourly[[6]]


#Plot for presentation to demonstrate a program

ggplot(component.estimation[[1]], aes(x = hr , y = X32.dtr12)) +
      geom_point(colour = "red", size = 3)+
      geom_line(colour = "red", size = 1.5)+
      scale_y_continuous(name = "Trait X", limits = c(0,.5), breaks = seq(0,.5,.1))+
      scale_x_continuous(name = "Time (hr)")+
      theme_classic(base_size = 20)+
      theme(legend.position = "none", legend.direction = "horizontal", panel.border = element_rect(colour = "black", fill=NA, size=1),  panel.grid.major.x = element_blank(), panel.grid.minor.x = element_blank(), panel.background = element_blank(), strip.background = element_blank(), strip.text = element_blank())




#Import experimental values
obs.df <- read.csv("Data/notfinished_obsvalues.csv")
obs.df$dtr <- as.factor(obs.df$dtr)

obs.bite <- subset(obs.df, trait == "bite.rate")
obs.EFD <- subset(obs.df, trait == "EFD")
obs.life.exp <- subset(obs.df, trait == "life.expect")
obs.life.egg <- subset(obs.df, trait == "life.eggs")

obs.r0 <- subset(obs.df, trait == "bite.rate")
obs.r0$obs <- r0_constant(obs.r0$temp)
obs.r0[is.na(obs.r0)] <- 0 


#Try an apply statement for ggplots - makes plots of estimated values
plots_mean <- lapply(trait.estimate, function(t){ #x is the dataframe in trait.estimate
  a <- ggplot(t, aes(x = temp, group = dtr, colour = dtr))+
           scale_colour_manual("Estimated:DTR", values = c("red", "pink", "black"))+
           geom_point(aes(y = means), size = 2.5, shape = 23, stroke = 1.5) +
           theme_classic(base_size = 20)
  
  return(a)
})

```

#Generate TPCS to be included in the plotes started above
```{r}
#BITE RATE

#Generate TPCs for observed and estimated DTR treatments - to be included in plots
#Estimate - DTR9
df.bite.estimate.dtr9 <- subset(trait.estimate[[1]],dtr==9)

nls_dtr9_briere_bite_start <- nls(means ~ c*temp*(temp-T0)*(Tm-temp)^(0.5), data= df.bite.estimate.dtr9, 
                start=list(c=0.001, T0=(-1), Tm=40.753))

temp <- df.bite.estimate.dtr9$temp
biterate <- df.bite.estimate.dtr9$means
bite_DF <-  cbind(temp, biterate)
bite_DF <- as.data.frame(bite_DF)

nls_dtr9_linear_bite <-nlsfit(bite_DF, model = 1)

nls_dtr9_quad_bite <- nlsfit(bite_DF, model = 2)

      #AIC values  
AIC <- c(nls_dtr9_linear_bite$Parameters[7,], nls_dtr9_quad_bite$Parameters[9,], AIC(nls_dtr9_briere_bite_start) )
Fit <- c("linear","quadratic", "briere")
bite_aic_table <- cbind(Fit, AIC)
bite_aic_table

nls_dtr9_briere_bite_fun.EST <-   function(x){ 0.0001425*x*(x-(-3.0895947))*(39.0878238 -x)^(0.5) }          

#Observed - DTR9
df.bite.observed.dtr9 <- subset(obs.df,dtr==9 & trait == "bite.rate")
nls_dtr9_briere_bite_start2 <- nls(obs ~ c*temp*(temp-T0)*(Tm-temp)^(0.5), data= df.bite.observed.dtr9, 
                start=list(c=0.001, T0=(-3), Tm=40.753))

temp <- df.bite.observed.dtr9$temp
biterate <- df.bite.observed.dtr9$obs
bite_DF <-  cbind(temp, biterate)
bite_DF <- as.data.frame(bite_DF)

nls_dtr9_linear_bite2 <-nlsfit(bite_DF, model = 1)
nls_dtr9_quad_bite2 <- nlsfit(bite_DF, model = 2)

      #AIC values  
AIC <- c(nls_dtr9_linear_bite2$Parameters[7,], nls_dtr9_quad_bite2$Parameters[9,], AIC(nls_dtr9_briere_bite_start2) )
Fit <- c("linear","quadratic", "briere")
bite_aic_table <- cbind(Fit, AIC)
bite_aic_table

nls_dtr9_briere_bite_fun.OBS <-   function(x){ 0.00021*x*(x-(0.8061))*(34.34 -x)^(0.5) }          

#Estimate - DTR12
df.bite.estimate.dtr12 <- subset(trait.estimate[[1]],dtr==12)

nls_dtr12_briere_bite_start <- nls(means ~ c*temp*(temp-T0)*(Tm-temp)^(0.5), data= df.bite.estimate.dtr12, 
                start=list(c=0.001, T0=(-1), Tm=40.753))

temp <- df.bite.estimate.dtr12$temp
biterate <- df.bite.estimate.dtr12$means
bite_DF <-  cbind(temp, biterate)
bite_DF <- as.data.frame(bite_DF)

nls_dtr12_linear_bite <-nlsfit(bite_DF, model = 1)

nls_dtr12_quad_bite <- nlsfit(bite_DF, model = 2)

      #AIC values  
AIC <- c(nls_dtr12_linear_bite$Parameters[7,], nls_dtr12_quad_bite$Parameters[9,], AIC(nls_dtr12_briere_bite_start) )
Fit <- c("linear","quadratic", "briere")
bite_aic_table <- cbind(Fit, AIC)
bite_aic_table

nls_dtr12_briere_bite_fun.EST <-   function(x){  0.0001449*x*(x-(-3.2485674))*(38.0638015 -x)^(0.5) }  


#Observed -DTR12

df.bite.observed.dtr12 <- subset(obs.df,dtr==12 & trait == "bite.rate")
nls_dtr12_briere_bite_start2 <- nls(obs ~ c*temp*(temp-T0)*(Tm-temp)^(0.5), data= df.bite.observed.dtr12, 
                start=list(c=0.001, T0=(-3), Tm=40.753))

temp <- df.bite.observed.dtr12$temp
biterate <- df.bite.observed.dtr12$obs
bite_DF12 <-  cbind(temp, biterate)
bite_DF12 <- as.data.frame(bite_DF12)

nls_dtr12_linear_bite2 <-nlsfit(bite_DF12, model = 1)
nls_dtr12_quad_bite2 <- nlsfit(bite_DF12, model = 2)

      #AIC values  
AIC <- c(nls_dtr12_linear_bite2$Parameters[7,], nls_dtr12_quad_bite2$Parameters[9,], AIC(nls_dtr12_briere_bite_start2) )
Fit <- c("linear","quadratic", "briere")
bite_aic_table <- cbind(Fit, AIC)
bite_aic_table

nls_dtr12_briere_bite_fun.OBS <-   function(x){ 0.00019*x*(x-(1.7703))*(34.21 -x)^(0.5) } 






###### Life expectancy #######
#Estimate - DTR9
df.lifeexpect.estimate.dtr9 <- subset(trait.estimate[[4]],dtr==9)

temp <- df.lifeexpect.estimate.dtr9$temp
lifeexpect <- df.lifeexpect.estimate.dtr9$means
lifeexpect_DF <-  cbind(temp, lifeexpect)
lifeexpect_DF <- as.data.frame(lifeexpect_DF)

#nls_dtr9_briere_lifeexpect_start <- nls(means ~ c*temp*(temp-T0)*(Tm-temp)^(0.5), data= df.lifeexpect.estimate.dtr9, start=list(c=0.000000135, T0= (-3880000), Tm=33))

library(thermPerf)

mFunction = function(x, params) {
  # params model parameters, a0, a1, a2
  a0 = params[["a0"]]
  a1 = params[["a1"]]
  a2 = params[["a2"]]
  return(a1 * x + a2 * x^2 + a0)
}
# 2) Name
mName = "Quadratic"
# 3) Formula
mFormula = y ~ a1 * x + a2 * x^2 + a0
# 4) Model parameters
mParams = c("a0", "a1", "a2")
# 5) List of starting values for the parameters
mStarting = list(a0 = 0, a1 = 1, a2 = 0.5)
# Create the customModel object
quadratic = buildModel(mFunction, mName, mFormula, mParams, mStarting)


models = getModelLibrary()[c( "briere1","linearFit")]
models[["quadratic"]] = quadratic
models$briere1$starting = list(a = .00001, t0 = 10, tmax = 40 )
Exfits = fitModels(models, temp, lifeexpect)
plot(Exfits, xlim = c(10, 40), ylim = c(0, 50), las = 1)
Exfitsweights = calculateAIC(Exfits)
plot(Exfitsweights)
#briere AIC: 31.66
#linear AIC: 31.71
#Quadratic: -307.58
nls_dtr9_linear_lifeexpect <-nlsfit(lifeexpect_DF, model = 1)

Exfits$quadratic$m$getPars()
nls_dtr9_lifeexpect_quad_fun.EST <-   function(x){ -2.922528 + 4.463632*x +(-0.118300)*x^2}          


#Observed - DTR9
df.lifeexpect.observed.dtr9 <- subset(obs.df,dtr==9 & trait == "life.expect")
nls_dtr9_briere_lifeexpect_start2 <- nls(obs ~ c*temp*(temp-T0)*(Tm-temp)^(0.5), data= df.lifeexpect.observed.dtr9, 
                start=list(c=0.1, T0= (2), Tm=40.753))

temp <- df.lifeexpect.observed.dtr9$temp
lifeexpect <- df.lifeexpect.observed.dtr9$obs
lifeexpect_DF <-  cbind(temp, lifeexpect)
lifeexpect_DF <- as.data.frame(lifeexpect_DF)

nls_dtr9_linear_lifeexpect2 <-nlsfit(lifeexpect_DF, model = 1)
nls_dtr9_quad_lifeexpect2 <- nlsfit(lifeexpect_DF, model = 2)

      #AIC values  
AIC <- c(nls_dtr9_linear_lifeexpect2$Parameters[7,], nls_dtr9_quad_lifeexpect2$Parameters[9,], AIC(nls_dtr9_briere_lifeexpect_start2) )
Fit <- c("linear","quadratic", "briere")
lifeexpect_aic_table <- cbind(Fit, AIC)
lifeexpect_aic_table

nls_dtr9_lifeexpect_quad_fun.OBS <-   function(x){ 28.2792 + 1.7576*x +(-0.0708)*x^2}        

#Estimate - DTR12
df.lifeexpect.estimate.dtr12 <- subset(trait.estimate[[4]],dtr==12)

#nls_dtr12_briere_lifeexpect_start <- nls(means ~ c*temp*(temp-T0)*(Tm-temp)^(0.5), data= df.lifeexpect.estimate.dtr12,   start=list(c=0.000000111, T0= (-4600000), Tm=33))



temp <- df.lifeexpect.estimate.dtr12$temp
lifeexpect <- df.lifeexpect.estimate.dtr12$means
lifeexpect_DF <-  cbind(temp, lifeexpect)
lifeexpect_DF <- as.data.frame(lifeexpect_DF)

models = getModelLibrary()[c( "briere1","linearFit")]
models[["quadratic"]] = quadratic
models$briere1$starting = list(a = .00001, t0 = 10, tmax = 40 )
Exfits = fitModels(models, temp, lifeexpect)
plot(Exfits, xlim = c(10, 40), ylim = c(0, 50), las = 1)
Exfitsweights = calculateAIC(Exfits)
plot(Exfitsweights)

nls_dtr12_linear_lifeexpect <-nlsfit(lifeexpect_DF, model = 1)
nls_dtr12_quad_lifeexpect <- nlsfit(lifeexpect_DF, model = 2)

      #AIC values  
AIC <- c(nls_dtr12_linear_bite$Parameters[7,], nls_dtr12_quad_bite$Parameters[9,] )
Fit <- c("linear","quadratic")
lifeexpect_aic_table <- cbind(Fit, AIC)
lifeexpect_aic_table

nls_dtr12_quad_lifeexpect_fun.EST <-   function(x){ -0.9298 + 4.1736*x +(-0.1114)*x^2} 


#Observed - DTR12
df.lifeexpect.observed.dtr12 <- subset(obs.df,dtr==12 & trait == "life.expect")
nls_dtr12_briere_lifeexpect_start2 <- nls(obs ~ c*temp*(temp-T0)*(Tm-temp)^(0.5), data= df.lifeexpect.observed.dtr12, 
                start=list(c=0.1, T0= (2), Tm=40.753))

temp <- df.lifeexpect.observed.dtr12$temp
lifeexpect <- df.lifeexpect.observed.dtr12$obs
lifeexpect_DF12 <-  cbind(temp, lifeexpect)
lifeexpect_DF12 <- as.data.frame(lifeexpect_DF12)

nls_dtr12_linear_lifeexpect2 <-nlsfit(lifeexpect_DF12, model = 1)
nls_dtr12_quad_lifeexpect2 <- nlsfit(lifeexpect_DF12, model = 2)

      #AIC values  
AIC <- c(nls_dtr12_linear_lifeexpect2$Parameters[7,], nls_dtr12_quad_lifeexpect2$Parameters[9,], AIC(nls_dtr12_briere_lifeexpect_start2) )
Fit <- c("linear","quadratic", "briere")
lifeexpect_aic_table <- cbind(Fit, AIC)
lifeexpect_aic_table

nls_dtr12_lifeexpect_briere_fun.OBS <-   function(x){(-0.02525)*x*(x-(34.36866))*(62.05932 -x)^(0.5) }        






###### Lifetime eggs #######
df.lifeegg.estimate.dtr9 <- subset(trait.estimate[[5]],dtr==9)

temp <- df.lifeegg.estimate.dtr9$temp
lifeegg <- df.lifeegg.estimate.dtr9$means
lifeegg_DF <-  cbind(temp, lifeegg)
lifeegg_DF <- as.data.frame(lifeegg_DF)

nls_dtr9_briere_lifeegg_start <- nls(means ~ c*temp*(temp-T0)*(Tm-temp)^(0.5), data= df.lifeegg.estimate.dtr9,  start=list(c=0.5, T0=(10.494188502), Tm=36.546948005))

nls_dtr9_linear_lifeegg <-nlsfit(lifeegg_DF, model = 1)
nls_dtr9_quad_lifeegg <- nlsfit(lifeegg_DF, model = 2)

      #AIC values  
AIC <- c(nls_dtr9_linear_lifeegg$Parameters[7,], nls_dtr9_quad_lifeegg$Parameters[9,], AIC(nls_dtr9_briere_lifeegg_start) )
Fit <- c("linear","quadratic", "briere")
lifeegg_aic_table <- cbind(Fit, AIC)
lifeegg_aic_table

nls_dtr9_quad_lifeegg_fun.EST <-   function(x){ -1151.8544 + 117.1790*x + (-2.3235)*x^2  }          

#Observed - DTR9
df.lifeegg.observed.dtr9 <- subset(obs.df,dtr==9 & trait == "life.eggs")
nls_dtr9_briere_lifeegg_start2 <- nls(obs ~ c*temp*(temp-T0)*(Tm-temp)^(0.5), data= df.lifeegg.observed.dtr9,start=list(c=1.0, T0=(5), Tm=33))

temp <- df.lifeegg.observed.dtr9$temp
lifeeggs <- df.lifeegg.observed.dtr9$obs
lifeeggs_DF <-  cbind(temp, lifeeggs)
lifeeggs_DF <- as.data.frame(lifeeggs_DF)

nls_dtr9_linear_lifeegg2 <-nlsfit(lifeeggs_DF, model = 1)
nls_dtr9_quad_lifeegg2 <- nlsfit(lifeeggs_DF, model = 2)

      #AIC values  
AIC <- c(nls_dtr9_linear_lifeegg2$Parameters[7,], nls_dtr9_quad_lifeegg2$Parameters[9,], AIC(nls_dtr9_briere_lifeegg_start2) )
Fit <- c("linear","quadratic", "briere")
lifeegg_aic_table <- cbind(Fit, AIC)
lifeegg_aic_table

nls_dtr9_quad_lifeegg_fun.OBS <-   function(x){ -1988.71+ 205.6601*x + (-4.2854)*x^2 }          

#Estimate - DTR12
df.lifeegg.estimate.dtr12 <- subset(trait.estimate[[5]],dtr==12)

temp <- df.lifeegg.estimate.dtr12$temp
lifeegg <- df.lifeegg.estimate.dtr12$means
lifeegg_DF <-  cbind(temp, lifeegg)
lifeegg_DF <- as.data.frame(lifeegg_DF)

nls_dtr12_briere_lifeegg_start <- nls(means ~ c*temp*(temp-T0)*(Tm-temp)^(0.5), data= df.lifeegg.estimate.dtr12, start=list(c=0.5, T0=(10.494188502), Tm=36.546948005))
nls_dtr12_linear_lifeegg <-nlsfit(lifeegg_DF, model = 1)
nls_dtr12_quad_lifeegg <- nlsfit(lifeegg_DF, model = 2)

      #AIC values  
AIC <- c(nls_dtr12_linear_lifeegg$Parameters[7,], nls_dtr12_quad_lifeegg$Parameters[9,], AIC(nls_dtr12_briere_lifeegg_start) )
Fit <- c("linear","quadratic", "briere")
lifeegg_aic_table <- cbind(Fit, AIC)
lifeegg_aic_table

nls_dtr12_quad_lifeegg_fun.EST <-   function(x){ -959.1674 + 99.9361*x + (-1.9801)*x^2 } 

#Observed - DTR12
df.lifeegg.observed.dtr12 <- subset(obs.df,dtr==12 & trait == "life.eggs")
nls_dtr12_briere_lifeegg_start2 <- nls(obs ~ c*temp*(temp-T0)*(Tm-temp)^(0.5), data= df.lifeegg.observed.dtr12,start=list(c=1.0, T0=(5), Tm=33))

temp <- df.lifeegg.observed.dtr12$temp
lifeeggs <- df.lifeegg.observed.dtr12$obs
lifeeggs_DF12 <-  cbind(temp, lifeeggs)
lifeeggs_DF12 <- as.data.frame(lifeeggs_DF12)

nls_dtr12_linear_lifeegg2 <-nlsfit(lifeeggs_DF12, model = 1)
nls_dtr12_quad_lifeegg2 <- nlsfit(lifeeggs_DF12, model = 2)

      #AIC values  
AIC <- c(nls_dtr12_linear_lifeegg2$Parameters[7,], nls_dtr12_quad_lifeegg2$Parameters[9,], AIC(nls_dtr12_briere_lifeegg_start2) )
Fit <- c("linear","quadratic", "briere")
lifeegg_aic_table <- cbind(Fit, AIC)
lifeegg_aic_table

nls_dtr12_quad_lifeegg_fun.OBS <-   function(x){ -2965.8809+ 298.1961*x + (-6.2444)*x^2 }          


```



```{r}
#Creates plots that adds the constant temperature TPC and the observed values
names(plots_mean) <- names

bite.plot <- plots_mean[[1]] + scale_fill_manual("Observed:DTR", values = c("black", "pink", "red")) + 
  stat_function(fun = bite.life_fun, colour = "black", linetype = "solid", size = 1.5) + 
  stat_function(fun = nls_dtr9_briere_bite_fun.EST, colour = "pink", linetype = "dashed", size = 1.5) +
  stat_function(fun = nls_dtr9_briere_bite_fun.OBS , colour = "pink", linetype = "solid", size = 1.5) +
  stat_function(fun = nls_dtr12_briere_bite_fun.EST, colour = "red", linetype = "dashed", size = 1.5) + 
  stat_function(fun = nls_dtr12_briere_bite_fun.OBS, colour = "red", linetype = "solid", size = 1.5)+ 
  geom_point(data = obs.bite, aes(y = obs, fill = dtr),shape = 21, colour = "black", size = 3, stroke = 1.5) +
  geom_point(aes(y = means), size = .1, shape = 23, stroke = 2.5) +
  geom_point(aes(y = means), colour = "black", size = 2.5, shape = 23, stroke = 1.5) + 
  scale_y_continuous(name = "Bite rate", limits = c(0,0.5))+ scale_x_continuous(name = "Temperature (C)", limits = c(10,40)) +  theme(legend.position = "none", legend.direction = "horizontal", panel.border = element_rect(colour = "black", fill=NA, size=.5),  panel.grid.major.x = element_blank(), panel.grid.minor.x = element_blank(), panel.background = element_blank(), strip.background = element_blank(), strip.text = element_blank())


#PPT TPC of bite rate only
ggplot(data.frame(x = c(0, 40)), aes(Temp = x))+ 
 theme_classic(base_size = 20)+
 stat_function(fun = bite.life_fun, colour = "black", linetype = "solid", size = 1.5) + 
  stat_function(fun = nls_dtr9_briere_bite_fun.OBS , colour = "pink", linetype = "solid", size = 1.5) +
  stat_function(fun = nls_dtr12_briere_bite_fun.OBS, colour = "red", linetype = "solid", size = 1.5)+ 
 scale_y_continuous(name = "Bite rate", limits = c(0,0.5))+ scale_x_continuous(name = "Temperature (C)", limits = c(10,40)) +     theme(legend.position = "none", legend.direction = "horizontal", panel.border = element_rect(colour = "black", fill=NA, size=.5),  panel.grid.major.x = element_blank(), panel.grid.minor.x = element_blank(), panel.background = element_blank(), strip.background = element_blank(), strip.text = element_blank())


E.plot <- plots_mean[[4]]  + scale_fill_manual("Observed:DTR", values = c("black", "pink", "red")) +
  stat_function(fun = life.expect_fun, colour = "black", linetype = "solid", size = 1.5) +
  stat_function(fun = nls_dtr9_lifeexpect_quad_fun.EST, colour = "pink", linetype = "dashed", size = 1.5)+
  stat_function(fun = nls_dtr9_lifeexpect_quad_fun.OBS, colour = "pink", linetype = "solid", size = 1.5)+
  stat_function(fun = nls_dtr12_quad_lifeexpect_fun.EST, colour = "red", linetype = "dashed", size = 1.5) +
  stat_function(fun = nls_dtr12_lifeexpect_briere_fun.OBS, colour = "red", linetype = "solid", size = 1.5)+
  geom_point(data = obs.life.exp, aes(y = obs, fill = dtr),shape = 21, colour = "black", size = 3, stroke = 1.5) +
  geom_point(aes(y = means), size = .1, shape = 23, stroke = 2.5) +
  geom_point(aes(y = means),colour = "black", size = 2.5, shape = 23, stroke = 1.5) +
  scale_y_continuous(name = "Lifespan", limits = c(0,53), breaks = seq(0,50,10))+
  scale_x_continuous(name = "Temperature (C)", limits = c(10,40)) +  theme(legend.position = "none", legend.direction = "horizontal", panel.border = element_rect(colour = "black", fill=NA, size=.5),  panel.grid.major.x = element_blank(), panel.grid.minor.x = element_blank(), panel.background = element_blank(), strip.background = element_blank(), strip.text = element_blank())


#PPT TPC of E only
ggplot(data.frame(x = c(0, 40)), aes(Temp = x))+ 
 theme_classic(base_size = 20)+
 stat_function(fun = life.expect_fun, colour = "black", linetype = "solid", size = 1.5) + 
  stat_function(fun = nls_dtr9_lifeexpect_quad_fun.OBS, colour = "pink", linetype = "solid", size = 1.5)+
   stat_function(fun = nls_dtr12_lifeexpect_briere_fun.OBS, colour = "red", linetype = "solid", size = 1.5)+
 scale_y_continuous(name = "Lifespan", limits = c(0,53), breaks = seq(0,50,10))+
   scale_x_continuous(name = "Temperature (C)", limits = c(10,40)) +
theme(legend.position = "none", legend.direction = "horizontal", panel.border = element_rect(colour = "black", fill=NA, size=.5),  panel.grid.major.x = element_blank(), panel.grid.minor.x = element_blank(), panel.background = element_blank(), strip.background = element_blank(), strip.text = element_blank())


B.plot <- plots_mean[[5]]  + scale_fill_manual("Observed:DTR", values = c("black", "pink", "red")) +
  stat_function(fun = life.eggs_fun, colour = "black", linetype = "solid", size = 1.5)  +
  stat_function(fun = nls_dtr9_quad_lifeegg_fun.EST, colour = "pink", linetype = "dashed", size = 1.5) +
  stat_function(fun = nls_dtr9_quad_lifeegg_fun.OBS, color = "pink", linetype = "solid", size = 1.5) +
  stat_function(fun = nls_dtr12_quad_lifeegg_fun.EST, colour = "red", linetype = "dashed", size = 1.5) +
  stat_function(fun = nls_dtr12_quad_lifeegg_fun.OBS, color = "red", linetype = "solid", size = 1.5) +
  geom_point(data = obs.life.egg , aes(y = obs, fill = dtr), shape = 21, colour = "black", size = 3, stroke = 1.5) + 
  geom_point(aes(y = means), size = .1, shape = 23, stroke = 2.5) +
  geom_point(aes(y = means),colour = "black", size = 2.5, shape = 23, stroke = 1.5) + scale_y_continuous(name = "Lifetime egg production", limits = c(0,750), breaks = seq(0,700,100))+ scale_x_continuous(name = "Temperature (C)", limits = c(10,40)) +  theme(legend.position = "none", legend.direction = "horizontal", panel.border = element_rect(colour = "black", fill=NA, size=.5),  panel.grid.major.x = element_blank(), panel.grid.minor.x = element_blank(), panel.background = element_blank(), strip.background = element_blank(), strip.text = element_blank())


#PPT TPC of B only
ggplot(data.frame(x = c(0, 40)), aes(Temp = x))+ 
 theme_classic(base_size = 20)+
 stat_function(fun = life.eggs_fun, colour = "black", linetype = "solid", size = 1.5) + 
  stat_function(fun = nls_dtr9_quad_lifeegg_fun.OBS, color = "pink", linetype = "solid", size = 1.5) +
  stat_function(fun = nls_dtr12_quad_lifeegg_fun.OBS, color = "red", linetype = "solid", size = 1.5) +
scale_y_continuous(name = "Lifetime egg production", limits = c(0,700), breaks = seq(0,700,100))+
   scale_x_continuous(name = "Temperature (C)", limits = c(10,40)) +
theme(legend.position = "none", legend.direction = "horizontal", panel.border = element_rect(colour = "black", fill=NA, size=.5),  panel.grid.major.x = element_blank(), panel.grid.minor.x = element_blank(), panel.background = element_blank(), strip.background = element_blank(), strip.text = element_blank())


plots_mean[[2]]  +  stat_function(fun = EFD.life_fun, colour = "black", linetype = "solid", size = 1.5)+  ggtitle( names[2])  + geom_point(data = obs.EFD, aes(y = obs),shape = 19, size = 4) 
plots_mean[[3]]  +  stat_function(fun = Mu.exp_fun, colour = "black", linetype = "solid", size = 1.5)+  ggtitle( names[3])  


plots_mean[[6]] + stat_function(fun = r0_constant, colour = "black", linetype = "solid", size =1.5) + geom_point(data = obs.r0, aes(y = obs), shape = 19, size =4 )

#TPC for just the other functions
PDR.plot <- ggplot( data.frame(x = c(15, 35))) + stat_function(fun = Shapiro_PDR, colour = "black", linetype = "solid", size = 1.5) +  scale_y_continuous(name = "Pathogen development rate", limits = c(0,.2), breaks = seq(0,.2,.05))+ scale_x_continuous(name = "Temperature (C)", limits = c(10,40)) + theme_classic(base_size = 20)+ theme(legend.position = "none", legend.direction = "horizontal", panel.border = element_rect(colour = "black", fill=NA, size=.5),  panel.grid.major.x = element_blank(), panel.grid.minor.x = element_blank(), panel.background = element_blank(), strip.background = element_blank(), strip.text = element_blank())

Vectcomp.plot <- ggplot( data.frame(x = c(15, 35))) + stat_function(fun = Shapiro_VectorComp , colour = "black", linetype = "solid", size = 1.5) +  scale_y_continuous(name = "Vector competence", limits = c(0,1), breaks = seq(0,1,.1))+ scale_x_continuous(name = "Temperature (C)", limits = c(10,40)) + theme_classic(base_size = 20) +  theme(legend.position = "none", legend.direction = "horizontal", panel.border = element_rect(colour = "black", fill=NA, size=.5),  panel.grid.major.x = element_blank(), panel.grid.minor.x = element_blank(), panel.background = element_blank(), strip.background = element_blank(), strip.text = element_blank())

MDR.plot <- ggplot( data.frame(x = c(15, 35))) + stat_function(fun = Mord_MDR, colour = "black", linetype = "solid", size = 1.5) +  scale_y_continuous(name = "Mosquito development rate", limits = c(0,.12), breaks = seq(0,.12,.03))+ scale_x_continuous(name = "Temperature (C)", limits = c(10,40)) + theme_classic(base_size = 20) +  theme(legend.position = "none", legend.direction = "horizontal", panel.border = element_rect(colour = "black", fill=NA, size=.5),  panel.grid.major.x = element_blank(), panel.grid.minor.x = element_blank(), panel.background = element_blank(), strip.background = element_blank(), strip.text = element_blank())

PEA.plot <- ggplot( data.frame(x = c(15, 35))) + stat_function(fun = Mord_PEA, colour = "black", linetype = "solid", size = 1.5) +  scale_y_continuous(name = "Probability of egg to adult survival", limits = c(0,1), breaks = seq(0,1,.2))+ scale_x_continuous(name = "Temperature (C)", limits = c(10,40))  + theme_classic(base_size = 20)+  theme(legend.position = "none", legend.direction = "horizontal", panel.border = element_rect(colour = "black", fill=NA, size=.5),  panel.grid.major.x = element_blank(), panel.grid.minor.x = element_blank(), panel.background = element_blank(), strip.background = element_blank(), strip.text = element_blank())

 
```

##Integrate component trait thermal performance into R0
##R0 with observed vs R0 with estimated at each treatment integrating life.eggs, life.expect, and bite rate with previous TPCS.

```{r}

###PICK UP HERE ADD IN THE NEW FUNCTION NAMES & DEFINE AN R0_OBS_DTR12

r0_obs_dtr9 <- function (x){(nls_dtr9_briere_bite_fun.OBS(x)^2 * Shapiro_VectorComp(x)* exp(-1/(Shapiro_PDR(x)*nls_dtr9_lifeexpect_quad_fun.OBS(x))) * nls_dtr9_quad_lifeegg_fun.OBS(x)* Mord_PEA(x) * Mord_MDR(x) * nls_dtr9_lifeexpect_quad_fun.OBS(x)^2) }
curve(r0_obs_dtr9, from = 15, to = 40)
  
r0_est_dtr9 <- function (x){(nls_dtr9_briere_bite_fun.EST(x)^2 * Shapiro_VectorComp(x)* exp(-1/(Shapiro_PDR(x)*nls_dtr9_lifeexpect_quad_fun.EST(x))) * nls_dtr9_quad_lifeegg_fun.EST(x)* Mord_PEA(x) * Mord_MDR(x) * nls_dtr9_lifeexpect_quad_fun.EST(x)^2) }
curve(r0_est_dtr9, from = 15, to = 40)
  
r0_est_dtr12 <- function (x){(nls_dtr12_briere_bite_fun.EST(x)^2 * Shapiro_VectorComp(x)* exp(-1/(Shapiro_PDR(x)*nls_dtr12_quad_lifeexpect_fun.EST(x))) *  nls_dtr12_quad_lifeegg_fun.EST(x)* Mord_PEA(x) * Mord_MDR(x) * nls_dtr12_quad_lifeexpect_fun.EST(x)^2) }
curve(r0_est_dtr12, from = 15, to = 40)

r0_obs_dtr12 <- function (x){(nls_dtr12_briere_bite_fun.OBS(x)^2 * Shapiro_VectorComp(x)* exp(-1/(Shapiro_PDR(x)*nls_dtr12_lifeexpect_briere_fun.OBS(x))) * nls_dtr12_quad_lifeegg_fun.OBS(x)* Mord_PEA(x) * Mord_MDR(x) * nls_dtr12_lifeexpect_briere_fun.OBS(x)^2) }
curve(r0_obs_dtr9, from = 15, to = 40)



r0.plot<-  ggplot(trait.estimate[[6]], aes(x= temp, group = dtr, colour = dtr) ) +
  scale_colour_manual("DTR", values = c("pink", "red"))+
  stat_function(fun = r0_constant, colour = "black", linetype = "solid", size = 1.5)+
  stat_function(fun = r0_obs_dtr9, colour = "pink", linetype = "solid", size = 1.5) +
  stat_function(fun = r0_est_dtr9, colour = "pink", linetype = "dashed", size = 1.5) +
  stat_function(fun = r0_est_dtr12, colour = "red", linetype = "dashed", size = 1.5) +
  stat_function(fun = r0_obs_dtr12, colour = "red", linetype  = "solid", size = 1.5) +
  geom_point(aes(y = means), size = .1, shape = 23, stroke = 2.5) +
  geom_point(aes(y = means),colour = "black", size = 2.5, shape = 23, stroke = 1.5) +
  scale_x_continuous(name = "Temperature  (C)",breaks = seq(15, 35, 5), limits=c(15, 35)) +
  scale_y_continuous(name = "Vectorial capacity",breaks = seq(0.0,4000,1000), limits = c(0.0,4000), expand = c(0,0.01)) +
  theme_classic(base_size = 20)+
  theme(legend.position = "none", legend.direction = "horizontal", panel.border = element_rect(colour = "black", fill=NA, size=.5),  panel.grid.major.x = element_blank(), panel.grid.minor.x = element_blank(), panel.background = element_blank(), strip.background = element_blank(), strip.text = element_blank())


#PPT r0 without the rate summation triangles
ggplot(data.frame(x = c(15, 35)), aes(Temp = x))+ 
 theme_classic(base_size = 20)+
  stat_function(fun = r0_constant, colour = "black", linetype = "solid", size = 1.5)+
  stat_function(fun = r0_obs_dtr9, colour = "pink", linetype = "solid", size = 1.5) +
  stat_function(fun = r0_est_dtr9, colour = "pink", linetype = "dashed", size = 1.5) +
  stat_function(fun = r0_est_dtr12, colour = "red", linetype = "dashed", size = 1.5) +
  stat_function(fun = r0_obs_dtr12, colour = "red", linetype  = "solid", size = 1.5) +
 scale_x_continuous(name = "Temperature  (C)",breaks = seq(15, 35, 5), limits=c(15, 35)) +
  scale_y_continuous(name = "Vectorial capacity",breaks = seq(0.0,4000,1000), limits = c(0.0,4000), expand = c(0,0.01)) +
theme(legend.position = "none", legend.direction = "horizontal", panel.border = element_rect(colour = "black", fill=NA, size=.5),  panel.grid.major.x = element_blank(), panel.grid.minor.x = element_blank(), panel.background = element_blank(), strip.background = element_blank(), strip.text = element_blank())
 
 #Integrating over R0 along falls to capture the complexity of the component traits
 
 #Making a panel of the 3 traits and VC
library(cowplot)

traits_panel <- plot_grid(bite.plot, E.plot, B.plot, r0.plot, labels=c("A", "B", "C", "D"), ncol = 1, nrow = 4, align = "hv", label_size = 20, vjust = 1.1)
#save_plot("traits_panel.png", traits_panel, ncol = 1.5, nrow = 4, base_aspect_ratio = 1.3)
 
traits_literature <- plot_grid(PDR.plot, Vectcomp.plot, MDR.plot, PEA.plot, labels=c("D", "E", "F", "G"), ncol = 1, nrow = 4, align = "hv", label_size = 20, vjust = 1.1)
#save_plot("traits_literature.png", traits_literature, ncol = 1.5, nrow = 4, base_aspect_ratio = 1.3)
 

  
```
#Estimating R0 at fluctuation from hourly intervals of R0 at constant (ie black line above)
```{r}


```

#Estimating component trait values at fluctuation using TPCs from constant temperature (1st GC, but not lifespan)

